Beam-Helicity and Beam-Charge Asymmetries \\ in the Bethe-Heitler and DVCS processes on \\ an Unpolarised Proton Target \underline{Paper Tag: PDVCS0607 (v2.0)} \underline{Drafting Committee: DC90 (\today)}

### 2011-09-13 background_and_systematic_dual.tex
Made Delias and wdns changes

The data sample in the exclusive region contains events involving both
production of real photons in which the proton remains intact and
events involving the excitation of the target proton to a $\Delta^+$
resonant state (``associated production''). The recoiling proton is not detected and
  the calorimeter resolution does not allow separation of the latter events from the rest of the data sample.
No systematic uncertainty is assigned for the contribution from these
events; they are treated as part of the signal. A Monte Carlo
calculation based on the parameterisation from ref.~\cite{Bra76} is
used to estimate the \red{fractional} contribution to the event sample from resonant
production in each kinematic bin. The results, called the associated fractions and labelled ``Assoc. fraction'', are shown in the last row of figure\red{s} in the results section. The method used to perform this estimation is described in detail in
ref.~\cite{Air08}.

>>>>

The data sample in the exclusive region contains events involving both
production of real photons in which the proton remains intact and
events involving the excitation of the target proton to a $\Delta^+$
resonant state (``associated production''). The recoiling proton is not detected and
  the calorimeter resolution does not allow separation of the latter events from the rest of the data sample.
No systematic uncertainty is assigned for the contribution from these
events; they are treated as part of the signal. A Monte Carlo
calculation based on the parameterisation from ref.~\cite{Bra76} is
used to estimate the fractional contribution to the event sample from resonant
production in each kinematic bin. The results, called the associated fractions and labelled ``Assoc. fraction'', are shown in the last row of figures in the results section. The method used to perform this estimation is described in detail in
ref.~\cite{Air08}.

### 2011-09-13 experiment_and_data.tex
Made Delias and wdns changes

The asymmetry amplitudes $A_{\textrm{LU,I/DVCS}}^{\sin(n\phi)}$ and
$A_{\textrm{C}}^{\cos(n\phi)}$ relate respectively to the Fourier
coefficients $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{I}$ from the interference and
DVCS terms in eqs.~\ref{e:alui}-\ref{e:ac}. The asymmetry amplitudes
may also be affected by the lepton propagators and the other
$\phi$-dependent terms in the denominators in
eqs.~\ref{e:alui}-\ref{e:ac}. The  $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{V}$ Fourier
coefficients depend on ``$\mathcal{C}$-functions'' \cite{{Bel02b}}, each of which is a
combination of complex Compton Form Factors (CFFs). A CFF is a
convolution of a GPD with a hard scattering kernel. Contributions of
GPDs to the cross section may be relatively suppressed by powers of
$1/Q$, the origin of which may be kinematic in nature or due to the
twist level of the GPD. Leading twist is twist-2. Typically, the
contribution of a twist-$n$ GPD, and hence the corresponding CFF, is
suppressed by $\mathcal{O}(1/Q^{n-2})$. Table~\ref{tab_amplitudes}
presents the asymmetry amplitudes extracted in this analysis and the
related Fourier coefficients, dominant $\mathcal{C}$-function, twist-level and level of suppression.

At H{\sc ermes} kinematic conditions, the leading-twist beam-helicity asymmetry amplitude is $A_{\textrm{LU,I}}^{\sin\phi}$ and the leading beam-charge asymmetry amplitude is $A^{\cos\phi}_{\textrm{C}}$. These asymmetry amplitudes, for an unpolarised target, are sensitive respectively to Fourier coefficients $s_{1,\textrm{unp}}^{\textrm{I}}$ and $c_{1,\textrm{unp}}^{\textrm{I}}$. Both of these Fourier coefficients have a dominant contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function:
\begin{eqnarray}
s_{1,\textrm{unp}}^{\textrm{I}} \approx&\,\,\,\,8\lambda ky(2-y)\mathfrak{Im}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\quad\textrm{and}&\label{eq:s1}
\\
 c_{1,\textrm{unp}}^{\textrm{I}} \approx&8k(2- 2y + y^{2})\mathfrak{Re}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\,,&\label{eq:c1}
\end{eqnarray}
where the definition of the kinematic factor $k$ is found
  in ref.\cite{Bel02b}. The $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function can be
written
\cite{Bel02b} 
\begin{equation}
 \mathcal{C}_{\textrm{unp}}^{\textrm{I}} = F_{1}\mathcal{H} + \frac{x_{\textrm{B}}}{2-x_{\textrm{B}}}(F_{1}+F_{2})\widetilde{\mathcal{H}} -\frac{t}{4M^{2}}F_{2}\mathcal{E},
\label{Eq_cunp}
\end{equation}
where $F_{1}$ and $F_{2}$ are respectively the Dirac and Pauli form
factors of the nucleon and $\mathcal{H}$, $\widetilde{\mathcal{H}}$ and
$\mathcal{E}$ are CFFs that relate respectively to the GPDs $H$,
$\widetilde{H}$ and $E$.  At H{\sc ermes} kinematic
conditions (where $x_{\textrm{B}}$ and $\frac{-t}{4M^2}$ are of order 0.1), the
contributions of CFFs $\widetilde{\mathcal{H}}$ and $\mathcal{E}$ can be
neglected in eq.~\ref{Eq_cunp} (in first approximation) with respect to $\mathcal{H}$ since they
are kinematically suppressed by an order of magnitude or more.
Hence, the behaviour of
$\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$ is determined by CFF $\mathcal{H}$
and therefore GPD $H$ can be constrained through
measurements of $A_{\textrm{LU,I}}^{\sin\phi}$ and $A^{\cos\phi}_{\textrm{C}}$.
The asymmetry amplitudes $A_{\textrm{LU},\textrm{I}}^{\sin(2\phi)}$,
$A^{\cos(0\phi)}_{\textrm{C}}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ also depend
upon this $\mathcal{C}$-function. However, the $A^{\cos(0\phi)}_{\textrm{C}}$ amplitude is kinematically suppressed compared to the $A^{\cos\phi}_{\textrm{C}}$ amplitude, and the  $A_{\textrm{LU,I}}^{\sin(2\phi)}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ amplitudes are higher-twist quantities (see Table~\ref{tab_amplitudes}).

The DVCS asymmetry amplitude $A^{\sin\phi}_{\textrm{LU,DVCS}}$ has a
contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{DVCS}}$-function,
which is bilinear in CFFs. However, this twist-3 amplitude is inherently small at HERMES kinematic conditions due to the size of the $s_{1}^{\textrm{DVCS}}$ Fourier coefficient compared to the contribution from the $c_{n}^{\textrm{BH}}$ coefficients in the denominator of Eq.~\ref{e:aludvcs}. As a result of the more complicated dependence on the CFFs and this suppression, it
is more difficult to constrain GPDs
via the measurement of $A^{\sin\phi}_{\textrm{LU,DVCS}}$ than from the
kinematically-unsuppressed leading-twist amplitudes.

>>>>

The asymmetry amplitudes $A_{\textrm{LU,I/DVCS}}^{\sin(n\phi)}$ and
$A_{\textrm{C}}^{\cos(n\phi)}$ relate respectively to the Fourier
coefficients $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{I}$ from the interference and
DVCS terms in eqs.~\ref{e:alui}-\ref{e:ac}. The asymmetry amplitudes
may also be affected by the lepton propagators and the other
$\phi$-dependent terms in the denominators in
eqs.~\ref{e:alui}-\ref{e:ac}. The  $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{V}$ Fourier
coefficients depend on ``$\mathcal{C}$-functions'' \cite{{Bel02b}}, each of which is a
combination of complex Compton Form Factors (CFFs). A CFF is a
convolution of a GPD with a hard scattering kernel. Contributions of
GPDs to the cross section may be relatively suppressed by powers of
$1/Q$, the origin of which may be kinematic in nature or due to the
twist-level of the GPD. Leading twist is twist-2. Typically, the
contribution of a twist-$n$ GPD, and hence the corresponding CFF, is
suppressed by $\mathcal{O}(1/Q^{n-2})$. Table~\ref{tab_amplitudes}
presents the asymmetry amplitudes extracted in this analysis and the
related Fourier coefficients, dominant $\mathcal{C}$-function, twist-level and level of suppression.

At H{\sc ermes} kinematic conditions, the leading twist beam-helicity asymmetry amplitude is $A_{\textrm{LU,I}}^{\sin\phi}$ and the leading beam-charge asymmetry amplitude is $A^{\cos\phi}_{\textrm{C}}$. These asymmetry amplitudes, for an unpolarised target, are sensitive respectively to Fourier coefficients $s_{1,\textrm{unp}}^{\textrm{I}}$ and $c_{1,\textrm{unp}}^{\textrm{I}}$. Both of these Fourier coefficients have a dominant contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function:
\begin{eqnarray}
s_{1,\textrm{unp}}^{\textrm{I}} \approx&\,\,\,\,8\lambda ky(2-y)\mathfrak{Im}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\quad\textrm{and}&\label{eq:s1}
\\
 c_{1,\textrm{unp}}^{\textrm{I}} \approx&8k(2- 2y + y^{2})\mathfrak{Re}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\,,&\label{eq:c1}
\end{eqnarray}
where the definition of the kinematic factor $k$ is found
  in ref.\cite{Bel02b}. The $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function can be
written
\cite{Bel02b} 
\begin{equation}
 \mathcal{C}_{\textrm{unp}}^{\textrm{I}} = F_{1}\mathcal{H} + \frac{x_{\textrm{B}}}{2-x_{\textrm{B}}}(F_{1}+F_{2})\widetilde{\mathcal{H}} -\frac{t}{4M^{2}}F_{2}\mathcal{E},
\label{Eq_cunp}
\end{equation}
where $F_{1}$ and $F_{2}$ are respectively the Dirac and Pauli form
factors of the nucleon and $\mathcal{H}$, $\widetilde{\mathcal{H}}$ and
$\mathcal{E}$ are CFFs that relate respectively to the GPDs $H$,
$\widetilde{H}$ and $E$.  At H{\sc ermes} kinematic
conditions (where $x_{\textrm{B}}$ and $\frac{-t}{4M^2}$ are of order 0.1), the
contributions of CFFs $\widetilde{\mathcal{H}}$ and $\mathcal{E}$ can be
neglected in eq.~\ref{Eq_cunp} (in first approximation) with respect to $\mathcal{H}$ since they
are kinematically suppressed by an order of magnitude or more.
Hence, the behaviour of
$\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$ is determined by CFF $\mathcal{H}$
and therefore GPD $H$ can be constrained through
measurements of $A_{\textrm{LU,I}}^{\sin\phi}$ and $A^{\cos\phi}_{\textrm{C}}$.
The asymmetry amplitudes $A_{\textrm{LU},\textrm{I}}^{\sin(2\phi)}$,
$A^{\cos(0\phi)}_{\textrm{C}}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ also depend
upon this $\mathcal{C}$-function. However, the
$A^{\cos(0\phi)}_{\textrm{C}}$ amplitude is kinematically suppressed
compared to the $A^{\cos\phi}_{\textrm{C}}$ amplitude, and the
$A_{\textrm{LU,I}}^{\sin(2\phi)}$ and $A^{\cos(2\phi)}_{\textrm{C}}$
amplitudes are higher twist quantities (see Table~\ref{tab_amplitudes}).

The DVCS asymmetry amplitude $A^{\sin\phi}_{\textrm{LU,DVCS}}$ has a
contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{DVCS}}$-function,
which is bilinear in CFFs. However, this twist-3 amplitude is inherently small at HERMES kinematic conditions due to the size of the $s_{1}^{\textrm{DVCS}}$ Fourier coefficient compared to the contribution from the $c_{n}^{\textrm{BH}}$ coefficients in the denominator of Eq.~\ref{e:aludvcs}. As a result of the more complicated dependence on the CFFs and this suppression, it
is more difficult to constrain GPDs
via the measurement of $A^{\sin\phi}_{\textrm{LU,DVCS}}$ than from the
kinematically-unsuppressed leading twist amplitudes.

### 2011-09-13 results_dual.tex
Made Delias and wdns changes

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated in the same procedure as \red{used to estimate the systematic uncertainties of} the beam-helicity asymmetries. The leading-twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regards to $A_{\textrm{C}}^{\cos\phi}$ can be found in table~\ref{tab_amplitudes}. The measured amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ but they have no discernible dependence on $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero and have no significant variation in value over the range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading-twist amplitudes.

>>>>

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated in the same procedure as used to estimate the systematic uncertainties of the beam-helicity asymmetries. The leading-twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regards to $A_{\textrm{C}}^{\cos\phi}$ can be found in table~\ref{tab_amplitudes}. The measured amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ but they have no discernible dependence on $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero and have no significant variation in value over the range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading-twist amplitudes.

### 2011-09-13 summary.tex
Made Delias and wdns changes

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a new unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in Ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 event sample. The asymmetry amplitudes extracted from the combined data set are the statistically most precise DVCS measurements presented by H{\sc ermes}. There is a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

>>>>

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a new unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 event sample. The asymmetry amplitudes extracted from the combined data set are the statistically most precise DVCS measurements presented by H{\sc ermes}. There is a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.
